The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-01-10 Paper.Rmd

We present our analysis of the green spruce aphid abundance data (see Data).
Independence was tested for each pair of the 10 sampling locations, and was 
rejected in every case.
<!--***DAN: Right, Shyamolina? I am assuming that because "Indep" does not 
appear anywhere in Table S11, this is correct, but please confirm or correct-->
Model selection among copulas was carried out separately for each pair of locations.
Best-fitting (lowest-AIC) copulas were non-normal
for the large majority of location pairs (Table \ref{SM-tab_count_bestcop}), and 
AIC weights for the best-fitting copulas (Table \ref{SM-tab_count_AICw}) 
were typically substiantially 
larger than those for the N copula (Table \ref{SM-tab_count_AICw_normal}), answering 
the first part of Q1 from the Introduction for these data.
Goodness of fit tests based on a Cramer-von Mises statistic 
(Table \ref{SM-tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also 
an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov 
statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of 
copula families we used was sufficiently broad that the copulas of all pairs of 
locations could be characterized by at least one of them, for these data.
Model-averaged lower- (Table \ref{SM-tab_count_LTdep_AICw}) and upper-tail
(Table \ref{SM-tab_count_UTdep_AICw}) dependence statistics were exclusively 
positive and were frequently substantially large, thereby differing 
from what an N copula would give (i.e., 0), and answering the second part of 
Q1 for these data. Apparently these aphid data have greater tail dependence 
(both lower and upper) than would be expected from an N-copula null hypothesis. 
Model-averaged lower-tail dependence statistics minus model-averaged 
upper-tail dependence statistics (Table \ref{SM-tab_count_LTmUT}) were positive 
for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), 
indicating that lower-tail dependence across space was much stronger than upper-tail 
dependence. Thus the spatial synchrony of rarity in green spruce aphid is stronger 
than the spatial synchrony of outbreak events. This also answers the third part of Q1.

>>>>

We present our analysis of the green spruce aphid abundance data (see Data).
Independence was tested for each pair of the 10 sampling locations, and was 
rejected in every case.
<!--***DAN: Right, Shyamolina? I am assuming that because "Indep" does not 
appear anywhere in Table S11, this is correct, but please confirm or correct.
***Shya: Yes, you are right, all the pairwise sampling locations for aphid data rejected independence test and show copula structures.-->
Model selection among copulas was carried out separately for each pair of locations.
Best-fitting (lowest-AIC) copulas were non-normal
for the large majority of location pairs (Table \ref{SM-tab_count_bestcop}), and 
AIC weights for the best-fitting copulas (Table \ref{SM-tab_count_AICw}) 
were typically substiantially 
larger than those for the N copula (Table \ref{SM-tab_count_AICw_normal}), answering 
the first part of Q1 from the Introduction for these data.
Goodness of fit tests based on a Cramer-von Mises statistic 
(Table \ref{SM-tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also 
an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov 
statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of 
copula families we used was sufficiently broad that the copulas of all pairs of 
locations could be characterized by at least one of them, for these data.
Model-averaged lower- (Table \ref{SM-tab_count_LTdep_AICw}) and upper-tail
(Table \ref{SM-tab_count_UTdep_AICw}) dependence statistics were exclusively 
positive and were frequently substantially large, thereby differing 
from what an N copula would give (i.e., 0), and answering the second part of 
Q1 for these data. Apparently these aphid data have greater tail dependence 
(both lower and upper) than would be expected from an N-copula null hypothesis. 
Model-averaged lower-tail dependence statistics minus model-averaged 
upper-tail dependence statistics (Table \ref{SM-tab_count_LTmUT}) were positive 
for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), 
indicating that lower-tail dependence across space was much stronger than upper-tail 
dependence. Thus the spatial synchrony of rarity in green spruce aphid is stronger 
than the spatial synchrony of outbreak events. This also answers the third part of Q1.
